---
title: "Error in using contrast"
author: "Wilfred Lee"
date: '2022-07-28'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(DESeq2)
library(dplyr)
```

# Make example data set with three conditions.

Create some random data:
```{r}
dds <- makeExampleDESeqDataSet(n=100, m=20, betaSD=1)
```

## Modify colData
 
```{r}
data_frame_colData <- dds@colData
colnames(data_frame_colData)[1] <- "treatment"
data_frame_colData$gender <- factor(sample(c("m", "f"), size = 20, replace = TRUE))
data_frame_colData$age <- factor(sample(c(10, 20, 30), size = 20, replace = TRUE))
data_frame_colData
```

```{r}
colData(dds) <- data_frame_colData
design(dds) <- ~ treatment     # Has to do this because the "condition" column no longer exists in colData.
dds
# Same as:
# dds@colData <- data_frame_colData
# dds@design <- ~ treatment   
```

## Setting reference levels

Set a custom reference level for each condition:
```{r}
dds$treatment <- relevel(dds$treatment, ref = "A")    #Not really needed for dds$treatment because already in the right order, but do this just in case.
dds$gender <- relevel(dds$gender, ref = "m")
dds$age <- relevel(dds$age, ref = "20")

```


## Age and treatment

To account for the effect of gender on treatment:
```{r message=FALSE}
dds_age_treatment <- dds
design(dds_age_treatment) <- ~age + treatment
dds_age_treatment <- DESeq(dds_age_treatment)
```

```{r}
results(dds_age_treatment, contrast = c("age", "30", "20"))
# Same as results(dds_age_treatment, contrast = list("age_30_vs_20", "age_10_vs_20"))
```
### Using the model matrix (not correct)
Another (appear to be incorrect) way to get the result, as given in [this tutorial](https://github.com/tavareshugo/tutorial_DESeq2_contrasts/) is:
```{r}
matrix_model <- model.matrix(design(dds_age_treatment), colData(dds_age_treatment))
array_for_contrast_10 <- colMeans(matrix_model[dds_age_treatment$age == "10", ])
array_for_contrast_20 <- colMeans(matrix_model[dds_age_treatment$age == "20", ])
array_for_contrast_30 <- colMeans(matrix_model[dds_age_treatment$age == "30", ])

array_for_contrast_A <- colMeans(matrix_model[dds_age_treatment$treatment == "A", ])
array_for_contrast_B <- colMeans(matrix_model[dds_age_treatment$treatment == "B", ])
```

```{r}
results(dds_age_treatment, contrast = array_for_contrast_30 - array_for_contrast_10 )
```
Unlike what was claimed in [the tutorial](https://github.com/tavareshugo/tutorial_DESeq2_contrasts/), the results does not match with `results(dds_age_treatment, contrast = c("age", "30", "10"))`. 

The two results below also disagree:
```{r}
results(dds_age_treatment, contrast = array_for_contrast_B - array_for_contrast_A)
results(dds_age_treatment)
```

